---
title: "Intro to Linear Regression"
author: "Larissa Neilson"
date: "11/14/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(palmerpenguins)
library(ggpubr)
library(broom)
```

# Rank-based test example (Mann-Whitney U)

## Create two sample vectors

```{r}
set.seed(1414)
gp_1 <- sample.int(20, size = 15, replace = TRUE)

set.seed(1424)
gp_2 <- sample.int(30, size = 15, replace = TRUE)
```

Is there evidence for a significant difference in ranks (medians) between the populations from which the vectors were drawn?

## Visualize with histograms

```{r}
hist(gp_1)

hist(gp_2)
```

### We might choose a rank-based test because:

1. Not clearly normally distributed from exploratory histograms
2. Somewhat small sample size (n = 15 for each)
3. I’ve decided that ranks (or, medians) are a more valuable metric to compare for these data.

## Perform Mann-Whitney U

"Is there a significant difference in ranks (medians) between these two vectors?"

```{r}
my_mwu <- wilcox.test(gp_1, gp_2)
```

P-value: 0.28

If the null hypothesis is true (these samples were drawn from populations with the same median), there is a probability of 0.28 that we could have found median values at least as different as ours by chance. There is not sufficient evidence to reject the null hypothesis of equal ranks (or medians) using a significance level of 0.05

# Simple Linear Regression

We'll use the penguins data set and explore the relationship between flipper length and body mass for penguins, including all 3 penguin species.

## Look at palmer penguins data

```{r}
ggplot(data = penguins, aes(x = flipper_length_mm, y = body_mass_g)) +
  geom_point()
```
1. Does it look like a linear relationship makes sense?
2. Do we have any concerns about modeling as a linear relationship?
3. Any notable outliers?
4. Initial thoughts about homoscedasticity (explored more later)?

## Model it

```{r}
penguin_lm <- lm(body_mass_g ~ flipper_length_mm, data = penguins)

summary(penguin_lm)
```

Both the intercept and flipper_length_mm coefficients are significantly different from zero (not super interesting).
The Multiple R2 value is 0.759 - meaning that 75.9% of variance in body mass is explained by flipper length

## Access model outputs

We can access the coefficients for the model using:

The slope is 49.69 (g / mm)
The y-intercept is -5780.83 (g)
The full equation is mass = 49.69*(flipper length) + (-5780.83)

Use `broom::tidy()` function to get the model outputs in nice data frame format

```{r}
penguin_lm_tidy <- tidy(penguin_lm)
```

Get the intercept

```{r}
penguin_int <- penguin_lm_tidy$estimate[1]
penguin_int
```

Get flipper length coefficient

```{r}
penguin_coef <- penguin_lm_tidy$estimate[2]
penguin_coef
```

Get some other model information (degrees of freedom, F-statistic, p-value, etc using `broom::glance()`

```{r}
penguin_lm_out <- glance(penguin_lm)
penguin_lm_out
```

We can use the results of both to write a statement about the model that will automatically update if anything about the model changes! Example:

"Simple linear regression was used to explore the relationship between penguin flipper length (mm) and body mass (g) across all three penguin species, and including both male and female penguins. A significant regression model was found ($\beta$ = `r round(penguin_coef,3)`, F(`r penguin_lm_out$df`,`r penguin_lm_out$df.residual`) = `r round(penguin_lm_out$statistic,1)`, p < 0.001) with an R^2^ of `r round(penguin_lm_out$r.squared,3)`."

## Explore model assumptions

Explore assumptinos for linear regression, some related to the residuals

Linearly related variables (CHECK - already looked & thought hard)
Normally distributed *residuals*
Homoscedasticity (constant residuals variance)
iid residuals (no serial correlation) - more often a concern in time series data

### Use `plot()` function onthe model to create four useful visualizations to consider assumptions!

```{r}
plot(penguin_lm)
```

What the four plots are:

The first one: fitted values vs. residuals
The second one: QQ-plot for residuals
The third one: another way of looking at fitted vs. residuals (these are just standardized residuals, but you can interpret it the same way)
The fourth one: Cook’s distance, a measure of “influence” or “leverage” that individual points have on the model - often considered a way to explore outliers.

## Visualize the model

Use `geom_smooth(method = "lm")` to add a linear model to an existing scatterplot

Use `stat_cor()` and/or `stat_regline_equation()` to add equation information directly to the plot panel, at an x- and y-position that you specify (and yes, you can mess with the digits & appearance here)

```{r}
ggplot(data = penguins, aes(x = flipper_length_mm, y = body_mass_g)) +
  geom_point(size = 2) +
  geom_smooth(method = "lm",
              color = "red",
              size = 0.5,
              fill = "gray10",
              alpha = 0.5) +
  theme_light() +
  stat_regline_equation(label.x = 180, label.y = 5700)
```
## Find Pearson's r for correlation

Typical guidelines:

r = 0: no correlation
r < |0.3|: weak correlation
r between |0.3| and |0.7|: moderate correlation
r > |0.7|: strong correlation

```{r}
penguins_cor <- cor.test(penguins$flipper_length_mm, penguins$body_mass_g)
```

Here, we see that there is a strong positive correlation between penguin flipper length and body mass (*r* = `r round(penguins_cor$estimate,2)`, t(`r penguins_cor$parameter`) = `r round(penguins_cor$statistic,2)`, p < 0.001).
